---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd and run `rmarkdown::render("README.Rmd")`. -->

```{r setup, include = FALSE}
# Cairo PNG + fig.bg: otherwise knitr defaults to an opaque white page and
# GitHub shows white behind "transparent" ggplot areas.
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 2.2,
  dpi = 120,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo", bg = "transparent"),
  fig.bg = "transparent"
)
dir.create("man/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r pkg, include = FALSE}
library(whopals)
library(ggplot2)
```

# whopals

This package brings the official [WHO Data Design Language](https://srhdteuwpubsa.z6.web.core.windows.net/gho/data/design-language/design-system/colors/) colours into R. Install it using:

```r
pak::pak("finlaycampbell/whopals")
```

Palettes are exposed as `whopals::pal_*` functions. Use the `theme` argument to specify a light or dark theme, `component` to specify the plot component such as text or base, and `variant` to select a palette variant where available. The available palettes are displayed below.

***

```{r plot-helper, include = FALSE}
# Dark canvas + white type: readable on GitHub; knitr uses transparent device bg.
plot_pal <- function(x) {
  lab <- names(x)
  if (is.null(lab)) {
    lab <- as.character(seq_along(x))
  }
  df <- data.frame(
    label = factor(lab, levels = lab),
    y = 1L,
    col = unname(x),
    stringsAsFactors = FALSE
  )
  ggplot(df, aes(x = label, y = y, fill = col)) +
    geom_tile(width = 0.92, height = 0.85) +
    scale_fill_identity() +
    coord_cartesian(expand = FALSE, clip = "off") +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, tag = NULL) +
    theme_void(base_size = 11) +
    theme(
      axis.text.x = element_text(
        angle = 35,
        hjust = 1,
        vjust = 1,
        color = "white",
        size = 9,
        margin = ggplot2::margin(t = 10, unit = "pt")
      ),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      axis.line = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_blank(),
      plot.subtitle = element_blank(),
      plot.caption = element_blank(),
      plot.tag = element_blank(),
      plot.margin = ggplot2::margin(6, 8, 22, 8),
      panel.background = element_rect(fill = "transparent", linewidth = 0),
      plot.background = element_rect(fill = "#0d1117", color = NA),
      panel.border = element_blank()
    )
}
```

**`pal_region()`**

```{r pal-region}
plot_pal(whopals::pal_region(theme = "dark"))
```

**`pal_trend()`**

```{r pal-trend}
plot_pal(whopals::pal_trend(theme = "dark"))
```

**`pal_gender()`**

```{r pal-gender}
plot_pal(whopals::pal_gender("dark"))
```

**`pal_category()`**

```{r pal-category}
plot_pal(whopals::pal_category(theme = "dark"))
```

**`pal_sequential("brand")`**

```{r pal-sequential-brand, fig.height = 2.2}
plot_pal(whopals::pal_sequential("brand"))
```

**`pal_sequential("complementary")`**

```{r pal-sequential-complementary, fig.height = 2.2}
plot_pal(whopals::pal_sequential("complementary"))
```

**`pal_sequential("colorful")`**

```{r pal-sequential-colorful, fig.height = 2.2}
plot_pal(whopals::pal_sequential("colorful"))
```

**`pal_diverging("base")`**

```{r pal-diverging-base, fig.height = 2.2}
plot_pal(whopals::pal_diverging("base", theme = "dark"))
```

**`pal_diverging("alt")`**

```{r pal-diverging-alt, fig.height = 2.2}
plot_pal(whopals::pal_diverging("alt", theme = "dark"))
```

**`pal_brand()`**

```{r pal-brand, fig.height = 1.5}
plot_pal(whopals::pal_brand(theme = "dark"))
```

**`pal_foreground()`**

```{r pal-foreground, fig.height = 1.5}
plot_pal(whopals::pal_foreground(theme = "dark"))
```

**`pal_background()`**

```{r pal-background, fig.height = 1.5}
plot_pal(whopals::pal_background(theme = "dark"))
```

**`pal_text()`**

```{r pal-text, fig.height = 1.5}
plot_pal(whopals::pal_text(theme = "dark"))
```

**`pal_functional()`**

```{r pal-functional, fig.height = 3.2}
plot_pal(whopals::pal_functional("dark"))
```

**`pal_selection()`**

```{r pal-selection, fig.height = 2.8}
plot_pal(whopals::pal_selection(theme = "dark"))
```

## Example column chart

The figures below show simple column charts of random values by WHO
region: **`pal_background()`** for the plot and panel fills,
**`pal_foreground()`** for axis titles and grid lines, **`pal_text()`** for
axis tick labels, **`pal_brand()`** for the title, and **`pal_region()`**
for bar fills. One figure uses **`theme = "dark"`**, the other
uses **`theme = "light"`**.

```{r example-column-data, echo=FALSE}
library(whopals)
library(ggplot2)
set.seed(42)
df_regions <- data.frame(
  who_region = c("afro", "amro", "emro", "euro", "searo", "wpro"),
  value = round(runif(6L, 8, 40), 1),
  stringsAsFactors = FALSE
)
```

```{r example-column-dark, echo=FALSE, fig.height = 3.6, fig.width = 8}
reg_dark <- whopals::pal_region(theme = "dark")
df_dark <- df_regions
df_dark$who_region <- factor(df_dark$who_region, levels = names(reg_dark))

ggplot(df_dark, aes(x = who_region, y = value, fill = who_region)) +
  geom_col(width = 0.72) +
  scale_fill_manual(values = reg_dark) +
  labs(title = "Illustrative values by WHO region", x = NULL, y = "Value") +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.background = element_rect(
      fill = unname(whopals::pal_background("base", "dark")),
      color = NA
    ),
    panel.background = element_rect(
      fill = unname(whopals::pal_background("selection", "dark")),
      color = NA
    ),
    plot.title = element_text(
      color = unname(whopals::pal_brand("base", "dark")),
      face = "bold",
      size = 14
    ),
    axis.title = element_text(
      color = unname(whopals::pal_foreground("base", "dark"))
    ),
    axis.text = element_text(
      color = unname(whopals::pal_text("base", "dark"))
    ),
    axis.text.x = element_text(angle = 28, hjust = 1, vjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha(unname(whopals::pal_foreground("weakest", "dark")), 0.4)
    )
  )
```

```{r example-column-light, echo=FALSE, fig.height = 3.6, fig.width = 8}
reg_light <- whopals::pal_region(theme = "light")
df_light <- df_regions
df_light$who_region <- factor(df_light$who_region, levels = names(reg_light))

ggplot(df_light, aes(x = who_region, y = value, fill = who_region)) +
  geom_col(width = 0.72) +
  scale_fill_manual(values = reg_light) +
  labs(title = "Illustrative values by WHO region", x = NULL, y = "Value") +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.background = element_rect(
      fill = unname(whopals::pal_background("base", "light")),
      color = NA
    ),
    panel.background = element_rect(
      fill = unname(whopals::pal_background("selection", "light")),
      color = NA
    ),
    plot.title = element_text(
      color = unname(whopals::pal_brand("base", "light")),
      face = "bold",
      size = 14
    ),
    axis.title = element_text(
      color = unname(whopals::pal_foreground("base", "light"))
    ),
    axis.text = element_text(
      color = unname(whopals::pal_text("base", "light"))
    ),
    axis.text.x = element_text(angle = 28, hjust = 1, vjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(
      color = alpha(unname(whopals::pal_foreground("weakest", "light")), 0.35)
    )
  )
```
