---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd and run `rmarkdown::render("README.Rmd")`. -->

```{r setup, include = FALSE}
# Cairo PNG + fig.bg: otherwise knitr defaults to an opaque white page and
# GitHub shows white behind "transparent" ggplot areas.
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 2.2,
  dpi = 120,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo", bg = "transparent"),
  fig.bg = "transparent"
)
dir.create("man/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r pkg, include = FALSE}
library(whopals)
library(ggplot2)
```

# whopals

This package brings the official WHO Data Design Language colour system into R as named hex values and small functions that return those palettes, so you can rely on the same published colours as on the WHO site without typing hex codes by hand. The specification is published by WHO at [WHO Data Design Language](https://srhdteuwpubsa.z6.web.core.windows.net/gho/data/design-language/design-system/colors/).

## Installation

```r
pak::pak("finlaycampbell/whopals")
```

After installing, every palette is exposed as a `whopals::pal_*` function. Use `theme` to choose a light or dark theme, `component` to specify the plot component (e.g. base vs text), and `variant` to select a palette variant where available.

```{r plot-helper, include = FALSE}
# Dark canvas + white type: readable on GitHub; knitr uses transparent device bg.
plot_pal <- function(x, title) {
  lab <- names(x)
  if (is.null(lab)) {
    lab <- as.character(seq_along(x))
  }
  df <- data.frame(
    label = factor(lab, levels = lab),
    y = 1L,
    col = unname(x),
    stringsAsFactors = FALSE
  )
  ggplot(df, aes(x = label, y = y, fill = col)) +
    geom_tile(width = 0.92, height = 0.85) +
    scale_fill_identity() +
    coord_cartesian(expand = FALSE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold", color = "white", size = 11),
      axis.text.x = element_text(
        angle = 35, hjust = 1, vjust = 1, color = "white", size = 9
      ),
      panel.background = element_rect(fill = "transparent", linewidth = 0),
      plot.background = element_rect(fill = "#0d1117", color = NA),
      panel.border = element_blank()
    )
}
```

### Palettes

Swatches use `theme = "dark"`. Figure titles repeat the R call including `theme = "dark"`. PNGs are saved with a transparent page background where supported; the plot area uses a dark face so labels stay readable on the light github.com page.

**`pal_theme()`**

* **theme**: light, dark

```{r pal-theme, fig.height = 2.8}
plot_pal(whopals::pal_theme("dark"), "whopals::pal_theme(theme = \"dark\")")
```

**`pal_category()`**

* **component**: base, stronger, text
* **theme**: light, dark

```{r pal-category}
plot_pal(
  whopals::pal_category(theme = "dark"),
  "whopals::pal_category(theme = \"dark\")"
)
```

**`pal_region()`**

* **component**: base, text
* **theme**: light, dark

```{r pal-region}
plot_pal(
  whopals::pal_region(theme = "dark"),
  "whopals::pal_region(theme = \"dark\")"
)
```

**`pal_sequential()`**

* **variant**: brand, complementary, colorful
* **component**: base, secondary, alt
* **theme**: light, dark

```{r pal-sequential-brand, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("brand", theme = "dark"),
  "whopals::pal_sequential(variant = \"brand\", theme = \"dark\")"
)
```

```{r pal-sequential-complementary, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("complementary", theme = "dark"),
  "whopals::pal_sequential(variant = \"complementary\", theme = \"dark\")"
)
```

```{r pal-sequential-colorful, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("colorful", theme = "dark"),
  "whopals::pal_sequential(variant = \"colorful\", theme = \"dark\")"
)
```

**`pal_diverging()`**

* **component**: base, alt
* **theme**: light, dark

```{r pal-diverging}
plot_pal(
  whopals::pal_diverging(theme = "dark"),
  "whopals::pal_diverging(theme = \"dark\")"
)
```

**`pal_functional()`**

* **theme**: light, dark

```{r pal-functional, fig.height = 3.2}
plot_pal(
  whopals::pal_functional("dark"),
  "whopals::pal_functional(theme = \"dark\")"
)
```

**`pal_gender()`**

* **theme**: light, dark

```{r pal-gender}
plot_pal(
  whopals::pal_gender("dark"),
  "whopals::pal_gender(theme = \"dark\")"
)
```

**`pal_trend()`**

* **component**: base, text, both
* **theme**: light, dark

```{r pal-trend}
plot_pal(
  whopals::pal_trend(theme = "dark"),
  "whopals::pal_trend(theme = \"dark\")"
)
```
