---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd and run `rmarkdown::render("README.Rmd")`. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 2.2,
  dpi = 120,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
dir.create("man/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r pkg, include = FALSE}
library(whopals)
library(ggplot2)
```

# whopals

**whopals** turns the colour definitions from the [WHO Data Design Language — Colours](https://srhdteuwpubsa.z6.web.core.windows.net/gho/data/design-language/design-system/colors/) into plain R: named hex vectors and small accessor functions so you can use the same themes, categories, regions, sequential and diverging scales, and functional colours in **ggplot2**, **highcharter**, or any other R graphics without hand-copying values from the site. Values are embedded from WHO design tokens (see `data-raw/build_whopals_colors.R` and `R/colors.R`).

## Installation

```r
pak::pak("finlaycampbell/whopals")
```

After `library(whopals)`, use `whopals::pal_theme()`, `whopals::pal_category()`, `whopals::pal_region()`, `whopals::pal_sequential()`, `whopals::pal_diverging()`, `whopals::pal_functional()`, `whopals::pal_gender()`, and `whopals::pal_trend()` (the `whopals::` prefix is optional once the package is attached). Defaults for each function match the plots below.

### Arguments by palette

**`pal_theme()`**

* **theme**: light, dark

**`pal_category()`**

* **component**: base, stronger, text
* **theme**: light, dark
* **include_other**: TRUE, FALSE

**`pal_region()`**

* **component**: base, text
* **theme**: light, dark

**`pal_sequential()`**

* **variant**: brand, complementary, colorful
* **component**: base, secondary (when **variant** is brand or complementary); base, alt (when **variant** is colorful)
* **theme**: light, dark
* **n**: NULL (discrete token stops) or an integer ≥ 2 (CIELAB interpolation length)

**`pal_diverging()`**

* **component**: base, alt
* **theme**: light, dark
* **n**: NULL (five stops) or an integer ≥ 2 (interpolated length)

**`pal_functional()`**

* **theme**: light, dark

**`pal_gender()`**

* **theme**: light, dark

**`pal_trend()`**

* **component**: base, text, both
* **theme**: light, dark

The plots below use **default** arguments only, e.g. `whopals::pal_gender()` matches `whopals::pal_gender(theme = "light")`.

```{r plot-helper, include = FALSE}
plot_pal <- function(x, title) {
  lab <- names(x)
  if (is.null(lab)) {
    lab <- as.character(seq_along(x))
  }
  df <- data.frame(
    label = factor(lab, levels = lab),
    y = 1L,
    col = unname(x),
    stringsAsFactors = FALSE
  )
  ggplot(df, aes(x = label, y = y, fill = col)) +
    geom_tile(width = 0.92, height = 0.85) +
    scale_fill_identity() +
    coord_cartesian(expand = FALSE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1)
    )
}
```

## `whopals::pal_theme()`

```{r pal-theme, fig.height = 2.8}
plot_pal(whopals::pal_theme(), "whopals::pal_theme()")
```

## `whopals::pal_category()`

```{r pal-category}
plot_pal(whopals::pal_category(), "whopals::pal_category()")
```

## `whopals::pal_region()`

```{r pal-region}
plot_pal(whopals::pal_region(), "whopals::pal_region()")
```

## `whopals::pal_sequential()`

```{r pal-sequential}
plot_pal(whopals::pal_sequential(), "whopals::pal_sequential()")
```

## `whopals::pal_diverging()`

```{r pal-diverging}
plot_pal(whopals::pal_diverging(), "whopals::pal_diverging()")
```

## `whopals::pal_functional()`

```{r pal-functional, fig.height = 3.2}
plot_pal(whopals::pal_functional(), "whopals::pal_functional()")
```

## `whopals::pal_gender()`

```{r pal-gender}
plot_pal(whopals::pal_gender(), "whopals::pal_gender()")
```

## `whopals::pal_trend()`

```{r pal-trend}
plot_pal(whopals::pal_trend(), "whopals::pal_trend()")
```
