---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd and run `rmarkdown::render("README.Rmd")`. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 2.2,
  dpi = 120,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
dir.create("man/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r pkg, include = FALSE}
library(whopals)
library(ggplot2)
```

# whopals

This package brings the official WHO Data Design Language colour system into R as named hex values and small functions that return those palettes, so you can rely on the same published colours as on the WHO site without typing hex codes by hand. The specification is published by WHO at [WHO Data Design Language](https://srhdteuwpubsa.z6.web.core.windows.net/gho/data/design-language/design-system/colors/).

## Installation

```r
pak::pak("finlaycampbell/whopals")
```

After installing, every palette is exposed as a `whopals::pal_*` function (for example `whopals::pal_theme()`, `whopals::pal_category()`, `whopals::pal_region()`, `whopals::pal_sequential()`, `whopals::pal_diverging()`, `whopals::pal_functional()`, `whopals::pal_gender()`, `whopals::pal_trend()`). You can always call them with the `whopals::` prefix; if you attach the package with `library(whopals)`, that prefix is optional.

```{r plot-helper, include = FALSE}
plot_pal <- function(x, title) {
  lab <- names(x)
  if (is.null(lab)) {
    lab <- as.character(seq_along(x))
  }
  df <- data.frame(
    label = factor(lab, levels = lab),
    y = 1L,
    col = unname(x),
    stringsAsFactors = FALSE
  )
  ggplot(df, aes(x = label, y = y, fill = col)) +
    geom_tile(width = 0.92, height = 0.85) +
    scale_fill_identity() +
    coord_cartesian(expand = FALSE) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1)
    )
}
```

### Options by palette

Examples use default arguments (including `theme = "light"`) unless the title states otherwise.

**`pal_theme()`**

* **theme**: light, dark

```{r pal-theme, fig.height = 2.8}
plot_pal(whopals::pal_theme(), "whopals::pal_theme()")
```

**`pal_category()`**

* **component**: base, stronger, text
* **theme**: light, dark

```{r pal-category}
plot_pal(whopals::pal_category(), "whopals::pal_category()")
```

**`pal_region()`**

* **component**: base, text
* **theme**: light, dark

```{r pal-region}
plot_pal(whopals::pal_region(), "whopals::pal_region()")
```

**`pal_sequential()`**

* **variant**: brand, complementary, colorful
* **component**: base, secondary, alt
* **theme**: light, dark

```{r pal-sequential-brand, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("brand"),
  'whopals::pal_sequential(variant = "brand")'
)
```

```{r pal-sequential-complementary, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("complementary"),
  'whopals::pal_sequential(variant = "complementary")'
)
```

```{r pal-sequential-colorful, fig.height = 2.2}
plot_pal(
  whopals::pal_sequential("colorful"),
  'whopals::pal_sequential(variant = "colorful")'
)
```

**`pal_diverging()`**

* **component**: base, alt
* **theme**: light, dark

```{r pal-diverging}
plot_pal(whopals::pal_diverging(), "whopals::pal_diverging()")
```

**`pal_functional()`**

* **theme**: light, dark

```{r pal-functional, fig.height = 3.2}
plot_pal(whopals::pal_functional(), "whopals::pal_functional()")
```

**`pal_gender()`**

* **theme**: light, dark

```{r pal-gender}
plot_pal(whopals::pal_gender(), "whopals::pal_gender()")
```

**`pal_trend()`**

* **component**: base, text, both
* **theme**: light, dark

```{r pal-trend}
plot_pal(whopals::pal_trend(), "whopals::pal_trend()")
```
