---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd and run `rmarkdown::render("README.Rmd")`. -->

```{r setup, include = FALSE}
# Cairo PNG + fig.bg: otherwise knitr defaults to an opaque white page and
# GitHub shows white behind "transparent" ggplot areas.
knitr::opts_chunk$set(
  echo = FALSE,
  fig.path = "man/figures/README-",
  fig.width = 8,
  fig.height = 2.2,
  dpi = 120,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dev.args = list(type = "cairo", bg = "transparent"),
  fig.bg = "transparent"
)
dir.create("man/figures", recursive = TRUE, showWarnings = FALSE)
```

```{r pkg, include = FALSE}
library(whopals)
library(ggplot2)
```

# whopals

This package brings the official [WHO Data Design Language](https://srhdteuwpubsa.z6.web.core.windows.net/gho/data/design-language/design-system/colors/) colours into R. Install it using:

```r
pak::pak("finlaycampbell/whopals")
```

Palettes are exposed as `whopals::pal_*` functions. Use the `theme` argument to specify a light or dark theme, `component` to specify the plot component such as text or base, and `variant` to select a palette variant where available. The available palettes are displayed below.

***

```{r plot-helper, include = FALSE}
# Dark canvas + white type: readable on GitHub; knitr uses transparent device bg.
plot_pal <- function(x) {
  lab <- names(x)
  if (is.null(lab)) {
    lab <- as.character(seq_along(x))
  }
  df <- data.frame(
    label = factor(lab, levels = lab),
    y = 1L,
    col = unname(x),
    stringsAsFactors = FALSE
  )
  ggplot(df, aes(x = label, y = y, fill = col)) +
    geom_tile(width = 0.92, height = 0.85) +
    scale_fill_identity() +
    coord_cartesian(expand = FALSE, clip = "off") +
    labs(x = NULL, y = NULL, title = NULL, subtitle = NULL, caption = NULL, tag = NULL) +
    theme_void(base_size = 11) +
    theme(
      axis.text.x = element_text(
        angle = 35,
        hjust = 1,
        vjust = 1,
        color = "white",
        size = 9,
        margin = ggplot2::margin(t = 10, unit = "pt")
      ),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      axis.line = element_blank(),
      panel.grid = element_blank(),
      plot.title = element_blank(),
      plot.subtitle = element_blank(),
      plot.caption = element_blank(),
      plot.tag = element_blank(),
      plot.margin = ggplot2::margin(6, 8, 22, 8),
      panel.background = element_rect(fill = "transparent", linewidth = 0),
      plot.background = element_rect(fill = "#0d1117", color = NA),
      panel.border = element_blank()
    )
}
```

**`pal_theme()`**

```{r pal-theme, fig.height = 2.8}
plot_pal(whopals::pal_theme("dark"))
```

**`pal_category()`**

```{r pal-category}
plot_pal(whopals::pal_category(theme = "dark"))
```

**`pal_region()`**

```{r pal-region}
plot_pal(whopals::pal_region(theme = "dark"))
```

**`pal_sequential("brand")`**

```{r pal-sequential-brand, fig.height = 2.2}
plot_pal(whopals::pal_sequential("brand"))
```

**`pal_sequential("complementary")`**

```{r pal-sequential-complementary, fig.height = 2.2}
plot_pal(whopals::pal_sequential("complementary"))
```

**`pal_sequential("colorful")`**

```{r pal-sequential-colorful, fig.height = 2.2}
plot_pal(whopals::pal_sequential("colorful"))
```

**`pal_diverging()`**

```{r pal-diverging}
plot_pal(whopals::pal_diverging(theme = "dark"))
```

**`pal_functional()`**

```{r pal-functional, fig.height = 3.2}
plot_pal(whopals::pal_functional("dark"))
```

**`pal_gender()`**

```{r pal-gender}
plot_pal(whopals::pal_gender("dark"))
```

**`pal_trend()`**

```{r pal-trend}
plot_pal(whopals::pal_trend(theme = "dark"))
```
